High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-10-16 Draft_V4046Sgr_mnras.tex
Update Draft_V4046Sgr_mnras.tex

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the {\tt uvmem}package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a pure $\chi^2$ model image. The resulting {\tt uvmem}image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at $\sim$35\,au into a shallow tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between
$\sim$10 and 17\,au.

@@

The ring width can be measured in polar coordinates by fitting 1-D Gaussians, thus obtaining a width and centroid at each azimuth. On average, we obtained a FWHM of 4.17$\pm$0.94\,au, and a stellocentric radius of 13.46$\pm$0.43\,au (See Fig. \ref{fig:polarring}). As the {\tt uvmem}model image has an effective angular resolution of $\sim$1/3 that of the natural-weighted beam ($0\farcs04 \,\times \, 0\farcs06$), we see that Ring13 is marginally resolved. After subtraction of the approximate {\tt uvmem} resolution ($\sim0\farcs01 \, \times \, \sim0\farcs02$), the ring width is $\sim$3.99$\pm$0.94\,au.

Repeating the optimization of the disk orientation, but this time aiming for Ring24 with a radial domain from 20\,au to 70\,au, we obtained a PA of 76.78$\pm$0.03\,$\degr$, with an inclination of 34.01$\pm$0.01\,$\degr$ and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24 both share the same orientation and centre, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24 are consistent within the errors, there are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ****CALCULATE THE CHI2 DIFFERENCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD****

@@

To search for a possible model that could explain the available data, we perform radiative transfer modelling with \textsc{radmc3d} \citep{Dullemond_2012}. Here we describe a 3D modelling framework for constructing disc structures given some parameters to characterise it. The model environment is similar to the one used by \citet{2018MNRAS.477.5104C} for DoAr 44. Through trial and error, we look for a set of values for the parameters that can define a model that correctly fits the available data.

@@

Assuming a three dimensional model in a spherical reference frame with coordinates ($r$, $\theta$, $\phi$), the gas density ($\rho_{\mathrm{gas}}$) distribution follows
\begin{equation}
  \rho_{\mathrm{gas}}(r,z) =\frac{\Sigma(r) \,\delta(r)}{\sqrt{2\pi} \,r H(r)}  \mathrm{exp}\left[-\frac{1}{2} \left(\frac{z}{r H(r)}\right)^2\right].
\end{equation}
Where $\delta(r)$ is a parameter that indicates the density drops in the gaps and cavities, $H(r)$ is the scale height profile and $\Sigma(r)$ been the surface density profile. $\Sigma(r)$ is defined by
\begin{equation}
  \Sigma(r) = \Sigma_\mathrm{c} \left(\frac{r}{R_\mathrm{c}}\right)^{-\gamma}  \, \mathrm{exp}\left[-\left(\frac{r}{R_\mathrm{c}}\right)^{2-\gamma}\right],
\end{equation}
where $R_c$ is a characteristic radius and $\gamma$ is the surface density gradient. A fixed $\gamma$ = 1 is used as it is a typical value for discs \citep{Andrews_2009,Andrews_2010}.

@@

The final structure of the parametric model goes as Fig. \ref{fig:profiles} shows. The inner radius of the model grid was set to 0.1\,au and an otter radius of 115\,au, large enough for the dust disc to become undetectable.

The dust begins at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au (Ring6),this ring is not present in the observations but was added for having a better SED fit. Then follows the two observed rings: the first goes from 12 to 19.5\,au (Ring13) and, with a 1\,au-wide gap between them, the outer ring (Ring24) goes from 20.5\,au and has its maximum density peak at 25\,au. The small-dust density then then fades as the radius increases.

@@

Fig. \ref{fig:SED} shows the SED constructed from the model compared to the photometry data available in the literature in \citet{Jensen_97} and online in \textsc{vizier} and \textsc{diana}, and also using spectrometry data from an archival \textit{Spitzer} IRS spectrum. From the similarity between the data and the resulting SED of the model we confirm that there has to be a small-grain population close to the stars down to 0.3\,au. The decision of employing the three ringed structure for the small dust grains relies on the fact that the SED needed a ring at a radius smaller than 10\,au to have a proper fit, this proposed ring (Ring6) has not been observed. The unexpected central emission in the 1.3\,mm continuum image could be part of this inner ring and, as it is asymmetric, the ring could be lopsided.

@@

%\begin{table}
% \caption{Model rings masses}
% \label{tab:masses}
% \begin{tabular}{lcc}
%  \hline
%  Ring & Location & Mass\\
%   & $\mathrm{au}$ & $M_{\earth}$ \\
%  \hline
%  First & 5 to 8 & 0.0007 \\
%  Second & 11 to 19 & 2.9 \\
%  Third & 20 to 80 & 43 \\
%  \hline
% \end{tabular}
%\end{table}

@@

New ALMA 1.3\,mm continuum imaging of the circumbinary disc around V4046\,Sgr reveals new information about its substructure. Using a radiative transfer model and decompositions in polar coordinates of the computed {\tt uvmem}model image, we measured and analysed the previously unseen thin inner ring and studied the small-dust population distribution.

>>>>

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the {\tt uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a $\chi^2$ model image with $\lambda$ = 0.1 for regularization. The resulting {\tt uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at $\sim$35\,au into a shallow tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between
$\sim$10 and 17\,au.

@@

The ring width can be measured in polar coordinates by fitting 1-D Gaussians, thus obtaining a width and centroid at each azimuth. On average, we obtained a FWHM of 4.17$\pm$0.94\,au, and a stellocentric radius of 13.46$\pm$0.43\,au (See Fig. \ref{fig:polarring}). As the {\tt uvmem} model image has an effective angular resolution of $\sim$1/3 that of the natural-weighted beam ($0\farcs04 \,\times \, 0\farcs06$), we see that Ring13 is marginally resolved. After subtraction of the approximate {\tt uvmem} resolution ($\sim0\farcs01 \, \times \, \sim0\farcs02$), the ring width is $\sim$3.99$\pm$0.94\,au.

Repeating the optimization of the disk orientation, but this time aiming for Ring24 with a radial domain from 20\,au to 70\,au, we obtained a PA of 76.78$\pm$0.03\,$\degr$, with an inclination of 34.01$\pm$0.01\,$\degr$ and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24 both share the same orientation and centre, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24 are consistent within the errors, there are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ***CALCULATE THE CHI2 DIFFERENCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD***

@@

To search for a possible model that could explain the available data, we perform radiative transfer modelling with \textsc{radmc3d} \citep{Dullemond_2012}. Here we describe a 3D modelling framework for constructing disc structures given some parameters to characterise it. The model environment is similar to the one used by \citet{2018MNRAS.477.5104C} for DoAr 44.

@@

Assuming a three dimensional model in a spherical reference frame with coordinates ($r$, $\theta$, $z$), the gas density ($\rho_{\mathrm{gas}}$) distribution follows
\begin{equation}
  \rho_{\mathrm{gas}}(r,z) =\frac{\Sigma(r) \,\delta(r)}{\sqrt{2\pi} \,r H(r)}  \mathrm{exp}\left[-\frac{1}{2} \left(\frac{z}{r H(r)}\right)^2\right].
\end{equation}
Where $\delta(r)$ is a parameter that indicates the density drops in the gaps and cavities, $H(r)$ is the scale height profile and $\Sigma(r)$ been the surface density profile. $\Sigma(r)$ is defined by
\begin{equation}
  \Sigma(r) = \Sigma_\mathrm{c} \left(\frac{r}{R_\mathrm{c}}\right)^{-\gamma}  \, \mathrm{exp}\left[-\left(\frac{r}{R_\mathrm{c}}\right)^{2-\gamma}\right],
\end{equation}
where $R_c$ is a characteristic radius and $\gamma$ is the surface density gradient. A fixed $\gamma$ = 1 is used as it is a typical value for discs \citep{Andrews_2009,Andrews_2010}.

@@

Through trial and error, we found a set of values for the parameters that define a model that correctly fits the available data. The final structure of the parametric model goes as Fig. \ref{fig:profiles} shows. The inner radius of the model grid was set to 0.1\,au and an otter radius of 115\,au, large enough for the dust disc to become undetectable.

The dust begins at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au, this ring (Ring6) is not present in the observations but was added for having a better SED fit. Then follows the two observed rings: the first goes from 12 to 19.5\,au (Ring13) and, with a 1\,au-wide gap between them, the outer ring (Ring24) goes from 20.5\,au and has its maximum density peak at 25\,au. The small-dust density then then fades as the radius increases.

@@

Fig. \ref{fig:SED} shows the SED constructed from the model compared to the photometry data available in the literature in \citet{Jensen_97} and online in \textsc{vizier} and \textsc{diana}, and also using spectrometry data from an archival \textit{Spitzer} IRS spectrum. From the similarity between the data and the resulting SED of the model we confirm that there has to be a small-grain population close to the stars down to 0.3\,au. The decision of employing the three ringed structure for the small dust grains relies on the fact that the SED needed a ring at a radius smaller than 10\,au to have a proper fit, this proposed ring (Ring6) has not been observed. The unexpected central emission in the 1.3\,mm continuum image could be part of this inner ring and, as it is asymmetric, the Ring6 could be lopsided.

@@

@@

New ALMA 1.3\,mm continuum imaging of the circumbinary disc around V4046\,Sgr reveals new information about its substructure. Using a radiative transfer model and decompositions in polar coordinates of the computed {\tt uvmem} model image, we measured and analysed the previously unseen thin inner ring and studied the small-dust population distribution.
